---
title: "MSDS 6371 Project"
author: "Halle Purdom & Taylor Bonar"
date: "4/4/2021"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

#loading libraries
library(ggplot2)
library(ggResidpanel)
library(naniar)
library(tidyr)
library(tidyverse)
library(plyr)
library(GGally)
library(ggthemes)
library(dplyr)
library(caret)
library(class)
library(e1071)
library(tm)
library(plotly)
library(scales)
library(sjPlot)
library(sjmisc)
library(sjlabelled)

train = read.csv("./train.csv", header = TRUE)
test = read.csv("./test.csv", header = TRUE)

```

# Initial Examination of the Datasets
```{r}
gg_miss_var(train)
```


# Analysis 1: How does the Square Footage of the Living Area Affect Sale Prices for the Houses in North Ames, Edwards, and Brookside?

## Step 1: Building and Fitting Our Initial Model:
$$
\hat{Sale Price} =  \hat{\beta_0} + \hat{\beta_1} * Living Area ft^2
$$
```{r}
# Filtering for desired neighborhoods into a separate dataframe
train_filtered = train %>% filter(Neighborhood == "NAmes" | Neighborhood == "Edwards" | Neighborhood == "BrkSide")

# Sales Prices' relation to square footage of the Living Area
## Create linear model of Response (SalePrice) to Explanatory Variables (GrLivArea)
fit1 = lm(SalePrice~GrLivArea, data=train_filtered)
## Find the overall Summary and Confidence Interval of our Linear Model
# Option 1 for auto-generating summary statistics:
tab_model(fit1, show.se = T, show.stat = T, show.aic = T, show.aicc = T,
          string.stat = "t-value", string.p = "p-value", string.se = "Std. Error", pred.labels = c("Intercept", "Sq. Ft. of Living Area"))
# Option 2
summary(fit1)
confint(fit1)
```
## Step 2: Checking Assumptions
### Examining Residual Plots & Influential Points
```{r Residual Plots, echo=F}
resid_panel(fit1, plots=c("resid","qq","ls","index","cookd","lev"))
```

### Addressing Assumptions
```{r}
## Generate Linear Regression Line
preds = predict(fit1)
train_filtered %>% ggplot(aes(y = SalePrice, x = GrLivArea)) +
  geom_point() + geom_smooth(method = "lm", formula = y~x) +
  ggtitle("Linear Regression Model: Living Area vs. Sale Price", "For Houses in North Ames, Edwards, and Brookside Neighborhoods")  +
  xlab("Square Footage of Living Area") +
  ylab("House Sale Price") +
  scale_y_continuous(labels = comma)

train_filtered %>% ggplot(aes(y = SalePrice, x = GrLivArea)) + 
  geom_point() + 
  geom_line(data = train_filtered, aes( x = GrLivArea, y = preds)) + 
  ggtitle("Linear Regression Model: Living Area vs. Sale Price", "For Houses in North Ames, Edwards, and Brookside Neighborhoods")  +
  xlab("Square Footage of Living Area") +
  ylab("House Sale Price") +
  scale_y_continuous(labels = comma)
```


# Performing Log Transforming on Data for GrLiveArea and SalePrice as New Columns
```{r}
# Create a new train set with log transformations w/ desired variables
train_log = train_filtered %>% data.frame(train_filtered$GrLivArea,train_filtered$SalePrice,train_filtered$Neighborhood)
# Translate names
names(train_log) <- c('GrLivArea', 'SalePrice','Neighborhood')
train_log$lGrLivArea = log(train_log$GrLivArea)
train_log$lSalePrice = log(train_log$SalePrice)
head(train_log)
```
## Log Transform only GrLivArea
```{r}
fit1 = lm(SalePrice~lGrLivArea, data=train_log)
summary(fit1)
confint(fit1)
par(mfrow = c(2, 2))
plot(fit1)
preds1 = predict(fit1)
train_log %>% ggplot(aes(y = SalePrice, x = lGrLivArea)) + geom_point() + geom_smooth(method = "lm")
train_log %>% ggplot(aes(y = SalePrice, x = lGrLivArea)) + geom_point() + geom_line(data = train_log, aes( x = lGrLivArea, y = preds1))
```

## Log Transform only Sale Price
```{r}
fit2 = lm(lSalePrice~GrLivArea, data=train_log)
summary(fit2)
confint(fit2)
par(mfrow = c(2, 2))
plot(fit2)
preds2 = predict(fit2)
train_log %>% ggplot(aes(y = lSalePrice, x = GrLivArea)) + geom_point() + geom_smooth(method = "lm")
train_log %>% ggplot(aes(y = lSalePrice, x = GrLivArea)) + geom_point() + geom_line(data = train_log, aes( x = GrLivArea, y = preds2))
```

## Log Transform Both GrLivArea and Sale Price
```{r}
fit3 = lm(lSalePrice~lGrLivArea, data=train_log)
summary(fit3)
confint(fit3)
par(mfrow = c(2, 2))
plot(fit3)
preds3 = predict(fit3)
train_log %>% ggplot(aes(y = lSalePrice, x = lGrLivArea)) + geom_point() + geom_smooth(method = "lm")
train_log %>% ggplot(aes(y = lSalePrice, x = lGrLivArea)) + geom_point() + geom_line(data = train_log, aes( x = lGrLivArea, y = preds3))
```
#separating based on neighborhood
```{r}
fit4 = lm(lSalePrice~lGrLivArea+ Neighborhood, data=train_log)
summary(fit4)
confint(fit4)
par(mfrow = c(2, 2))
plot(fit4)
preds4 = predict(fit4)
train_log %>% ggplot(aes(y = lSalePrice, x = lGrLivArea, color=Neighborhood)) + geom_point() + geom_smooth(method = "lm")
train_log %>% ggplot(aes(y = lSalePrice, x = lGrLivArea, color=Neighborhood)) + geom_point() + geom_line(data = train_log, aes(group=Neighborhood, x = lGrLivArea, y = preds4))
```



#selection models library
```{r}
library(olsrr)
library(cowplot)
library(ggplot2)
library(IDPmisc)
```

#only int and numeric type
```{r}
train_q2 = data.frame(train)

train_q2$lGrLivArea = log(train_q2$GrLivArea)
train_q2$lSalePrice = log(train_q2$SalePrice)

for(i in 1:ncol(train_q2)) {      
  if(typeof(train_q2[ , i]) == "integer" | typeof(train_q2[ , i]) == "double"){
    cat('"',colnames(train_q2)[i],'"',"= train_q2$",sep="", colnames(train_q2)[i], ", ")
  }
}

train_q2 <- data.frame("Id"= train_q2$Id, "MSSubClass"= train_q2$MSSubClass, "LotFrontage"= train_q2$LotFrontage, "LotArea"= train_q2$LotArea, "OverallQual"= train_q2$OverallQual, "OverallCond"= train_q2$OverallCond, "YearBuilt"= train_q2$YearBuilt, "YearRemodAdd"= train_q2$YearRemodAdd, "MasVnrArea"= train_q2$MasVnrArea, "BsmtFinSF1"= train_q2$BsmtFinSF1, "BsmtFinSF2"= train_q2$BsmtFinSF2, "BsmtUnfSF"= train_q2$BsmtUnfSF, "TotalBsmtSF"= train_q2$TotalBsmtSF, "X1stFlrSF"= train_q2$X1stFlrSF, "X2ndFlrSF"= train_q2$X2ndFlrSF, "LowQualFinSF"= train_q2$LowQualFinSF, "BsmtFullBath"= train_q2$BsmtFullBath, "BsmtHalfBath"= train_q2$BsmtHalfBath, "FullBath"= train_q2$FullBath, "HalfBath"= train_q2$HalfBath, "BedroomAbvGr"= train_q2$BedroomAbvGr, "KitchenAbvGr"= train_q2$KitchenAbvGr, "TotRmsAbvGrd"= train_q2$TotRmsAbvGrd, "Fireplaces"= train_q2$Fireplaces, "GarageYrBlt"= train_q2$GarageYrBlt, "GarageCars"= train_q2$GarageCars, "GarageArea"= train_q2$GarageArea, "WoodDeckSF"= train_q2$WoodDeckSF, "OpenPorchSF"= train_q2$OpenPorchSF, "EnclosedPorch"= train_q2$EnclosedPorch, "X3SsnPorch"= train_q2$X3SsnPorch, "ScreenPorch"= train_q2$ScreenPorch, "PoolArea"= train_q2$PoolArea, "MiscVal"= train_q2$MiscVal, "MoSold"= train_q2$MoSold, "YrSold"= train_q2$YrSold, "lGrLivArea"= train_q2$lGrLivArea, "lSalePrice"= train_q2$lSalePrice) #removed sale price and grlivarea
```

#adding logged variables to train dataset
```{r}

#TotalBsmtSF 
train_q2$lTotalBsmtSF = log(train_q2$TotalBsmtSF)
lTotalBsmtSF <- train_q2%>%ggplot(aes(x=lTotalBsmtSF, y=lSalePrice))+geom_point()+geom_smooth(method="lm")
TotalBsmtSF <- train_q2%>%ggplot(aes(x=TotalBsmtSF, y=lSalePrice))+geom_point()+geom_smooth(method="lm")
plot_grid(TotalBsmtSF,lTotalBsmtSF, labels = "AUTO")

#X1stFlrSF 
train_q2$lX1stFlrSF = log(train_q2$X1stFlrSF)
lX1stFlrSF <- train_q2%>%ggplot(aes(x=lX1stFlrSF, y=lSalePrice))+geom_point()+geom_smooth(method="lm")
X1stFlrSF <- train_q2%>%ggplot(aes(x=X1stFlrSF, y=lSalePrice))+geom_point()+geom_smooth(method="lm")
plot_grid(X1stFlrSF,lX1stFlrSF, labels = "AUTO")

#X2ndFlrSF
train_q2$lX2ndFlrSF = log(train_q2$X2ndFlrSF)
lX2ndFlrSF <- train_q2%>%ggplot(aes(x=lX2ndFlrSF, y=lSalePrice))+geom_point()+geom_smooth(method="lm")
X2ndFlrSF <- train_q2%>%ggplot(aes(x=X2ndFlrSF, y=lSalePrice))+geom_point()+geom_smooth(method="lm")
plot_grid(X2ndFlrSF,lX2ndFlrSF, labels = "AUTO")
```

#missing vars train_q2
```{r}
gg_miss_var(train_q2)
#lot frontage, garage year built, MasVnrArea
```

#remove lot frontage, garage year built, and masvnrarea
```{r}

train_q2 <- data.frame("Id"= train_q2$Id, "MSSubClass"= train_q2$MSSubClass, "LotArea"= train_q2$LotArea, "OverallQual"= train_q2$OverallQual, "OverallCond"= train_q2$OverallCond, "YearBuilt"= train_q2$YearBuilt, "YearRemodAdd"= train_q2$YearRemodAdd, "BsmtFinSF1"= train_q2$BsmtFinSF1, "BsmtFinSF2"= train_q2$BsmtFinSF2, "BsmtUnfSF"= train_q2$BsmtUnfSF, "TotalBsmtSF"= train_q2$TotalBsmtSF, "X1stFlrSF"= train_q2$X1stFlrSF, "X2ndFlrSF"= train_q2$X2ndFlrSF, "LowQualFinSF"= train_q2$LowQualFinSF, "BsmtFullBath"= train_q2$BsmtFullBath, "BsmtHalfBath"= train_q2$BsmtHalfBath, "FullBath"= train_q2$FullBath, "HalfBath"= train_q2$HalfBath, "BedroomAbvGr"= train_q2$BedroomAbvGr, "KitchenAbvGr"= train_q2$KitchenAbvGr, "TotRmsAbvGrd"= train_q2$TotRmsAbvGrd, "Fireplaces"= train_q2$Fireplaces, "GarageCars"= train_q2$GarageCars, "GarageArea"= train_q2$GarageArea, "WoodDeckSF"= train_q2$WoodDeckSF, "OpenPorchSF"= train_q2$OpenPorchSF, "EnclosedPorch"= train_q2$EnclosedPorch, "X3SsnPorch"= train_q2$X3SsnPorch, "ScreenPorch"= train_q2$ScreenPorch, "PoolArea"= train_q2$PoolArea, "MiscVal"= train_q2$MiscVal, "MoSold"= train_q2$MoSold, "YrSold"= train_q2$YrSold, "lGrLivArea"= train_q2$lGrLivArea, "lSalePrice"= train_q2$lSalePrice, "lX2ndFlrSF"= train_q2$lX2ndFlrSF, "lTotalBsmtSF"= train_q2$lTotalBsmtSF, "lX1stFlrSF"= train_q2$lX1stFlrSF)
```


#Error in lm.fit(x, y, offset = offset, singular.ok = singular.ok, ...) : NA/NaN/Inf in 'x'
```{r}
train_q2 <- NaRV.omit(train_q2)
```

#forward
```{r}
model_f <- lm(lSalePrice ~., data = train_q2)
f <-ols_step_forward_p(model_f, details=TRUE)
plot(f)
f
```

#backward
```{r}
model_b <- lm(lSalePrice ~ ., data = train_q2)
b <-ols_step_backward_p(model_b, details = TRUE)
plot(b)
b
```


#stepwwise
```{r}
model_s <- lm(lSalePrice ~ ., data = train_q2)
s <-ols_step_both_p(model_s, details = TRUE)
plot(s)
s
```

# Appendix
```{r ref.label="Residual Plots", eval=F}

