---
title: "MSDS 6371 Project"
author: "Halle"
date: "4/4/2021"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```
#loading libraries
```{r}
library(ggplot2)
library(naniar)
library(tidyr)
library(tidyverse)
library(plyr)
library(GGally)
library(ggthemes)
library(dplyr)
library(caret)
library(class)
library(e1071)
library(tm)
library(plotly)
```

#getting/examining the data
```{r}
train = read.csv("/Users/hallepurdom/Desktop/SMU/Classes/Stats Project/house-prices-advanced-regression-techniques/train.csv", header = TRUE)
test = read.csv("/Users/hallepurdom/Desktop/SMU/Classes/Stats Project/house-prices-advanced-regression-techniques/test.csv", header = TRUE)

gg_miss_var(train)
```

#filter for NAmes, Edwards and BrkSide neighborhoods
```{r}
train_filtered = train %>% filter(train$Neighborhood == "NAmes"|train$Neighborhood =="Edwards"|train$Neighborhood =="BrkSide")
```

#salesprice relation to square footage
```{r}
fit = lm(SalePrice~GrLivArea, data=train_filtered)
summary(fit)
confint(fit)
par(mfrow = c(2, 2))
plot(fit)
preds = predict(fit)
train_filtered %>% ggplot(aes(y = SalePrice, x = GrLivArea)) + geom_point() + geom_smooth(method = "lm")
train_filtered %>% ggplot(aes(y = SalePrice, x = GrLivArea)) + geom_point() + geom_line(data = train_filtered, aes( x = GrLivArea, y = preds))
```

#log transforming data
```{r}
train_log = data.frame(train_filtered$GrLivArea,train_filtered$SalePrice,train_filtered$Neighborhood)
names(train_log) <- c('GrLivArea', 'SalePrice','Neighborhood')
train_log$lGrLivArea = log(train_log$GrLivArea)
train_log$lSalePrice = log(train_log$SalePrice)
head(train_log)
```
#different log transform options -- 1
```{r}
#1
fit1 = lm(SalePrice~lGrLivArea, data=train_log)
summary(fit1)
confint(fit1)
par(mfrow = c(2, 2))
plot(fit1)
preds1 = predict(fit1)
train_log %>% ggplot(aes(y = SalePrice, x = lGrLivArea)) + geom_point() + geom_smooth(method = "lm")
train_log %>% ggplot(aes(y = SalePrice, x = lGrLivArea)) + geom_point() + geom_line(data = train_log, aes( x = lGrLivArea, y = preds1))
```

#different log transform options -- 2
```{r}
fit2 = lm(lSalePrice~GrLivArea, data=train_log)
summary(fit2)
confint(fit2)
par(mfrow = c(2, 2))
plot(fit2)
preds2 = predict(fit2)
train_log %>% ggplot(aes(y = lSalePrice, x = GrLivArea)) + geom_point() + geom_smooth(method = "lm")
train_log %>% ggplot(aes(y = lSalePrice, x = GrLivArea)) + geom_point() + geom_line(data = train_log, aes( x = GrLivArea, y = preds2))
```

#different log transform options -- 3
```{r}
fit3 = lm(lSalePrice~lGrLivArea, data=train_log)
summary(fit3)
confint(fit3)
par(mfrow = c(2, 2))
plot(fit3)
preds3 = predict(fit3)
train_log %>% ggplot(aes(y = lSalePrice, x = lGrLivArea)) + geom_point() + geom_smooth(method = "lm")
train_log %>% ggplot(aes(y = lSalePrice, x = lGrLivArea)) + geom_point() + geom_line(data = train_log, aes( x = lGrLivArea, y = preds3))
```
#separating based on neighborhood
```{r}
fit4 = lm(lSalePrice~lGrLivArea+ Neighborhood, data=train_log)
summary(fit4)
confint(fit4)
par(mfrow = c(2, 2))
plot(fit4)
preds4 = predict(fit4)
train_log %>% ggplot(aes(y = lSalePrice, x = lGrLivArea, color=Neighborhood)) + geom_point() + geom_smooth(method = "lm")
train_log %>% ggplot(aes(y = lSalePrice, x = lGrLivArea, color=Neighborhood)) + geom_point() + geom_line(data = train_log, aes(group=Neighborhood, x = lGrLivArea, y = preds4))
```



#selection models library
```{r}
library(olsrr)
library(cowplot)
library(ggplot2)
library(IDPmisc)
```

#only int and numeric type
```{r}
train_q2 = data.frame(train)

train_q2$lGrLivArea = log(train_q2$GrLivArea)
train_q2$lSalePrice = log(train_q2$SalePrice)

for(i in 1:ncol(train_q2)) {      
  if(typeof(train_q2[ , i]) == "integer" | typeof(train_q2[ , i]) == "double"){
    cat('"',colnames(train_q2)[i],'"',"= train_q2$",sep="", colnames(train_q2)[i], ", ")
  }
}

train_q2 <- data.frame("Id"= train_q2$Id, "MSSubClass"= train_q2$MSSubClass, "LotFrontage"= train_q2$LotFrontage, "LotArea"= train_q2$LotArea, "OverallQual"= train_q2$OverallQual, "OverallCond"= train_q2$OverallCond, "YearBuilt"= train_q2$YearBuilt, "YearRemodAdd"= train_q2$YearRemodAdd, "MasVnrArea"= train_q2$MasVnrArea, "BsmtFinSF1"= train_q2$BsmtFinSF1, "BsmtFinSF2"= train_q2$BsmtFinSF2, "BsmtUnfSF"= train_q2$BsmtUnfSF, "TotalBsmtSF"= train_q2$TotalBsmtSF, "X1stFlrSF"= train_q2$X1stFlrSF, "X2ndFlrSF"= train_q2$X2ndFlrSF, "LowQualFinSF"= train_q2$LowQualFinSF, "BsmtFullBath"= train_q2$BsmtFullBath, "BsmtHalfBath"= train_q2$BsmtHalfBath, "FullBath"= train_q2$FullBath, "HalfBath"= train_q2$HalfBath, "BedroomAbvGr"= train_q2$BedroomAbvGr, "KitchenAbvGr"= train_q2$KitchenAbvGr, "TotRmsAbvGrd"= train_q2$TotRmsAbvGrd, "Fireplaces"= train_q2$Fireplaces, "GarageYrBlt"= train_q2$GarageYrBlt, "GarageCars"= train_q2$GarageCars, "GarageArea"= train_q2$GarageArea, "WoodDeckSF"= train_q2$WoodDeckSF, "OpenPorchSF"= train_q2$OpenPorchSF, "EnclosedPorch"= train_q2$EnclosedPorch, "X3SsnPorch"= train_q2$X3SsnPorch, "ScreenPorch"= train_q2$ScreenPorch, "PoolArea"= train_q2$PoolArea, "MiscVal"= train_q2$MiscVal, "MoSold"= train_q2$MoSold, "YrSold"= train_q2$YrSold, "lGrLivArea"= train_q2$lGrLivArea, "lSalePrice"= train_q2$lSalePrice) #removed sale price and grlivarea
```

#adding logged variables to train dataset
```{r}

#TotalBsmtSF 
train_q2$lTotalBsmtSF = log(train_q2$TotalBsmtSF)
lTotalBsmtSF <- train_q2%>%ggplot(aes(x=lTotalBsmtSF, y=lSalePrice))+geom_point()+geom_smooth(method="lm")
TotalBsmtSF <- train_q2%>%ggplot(aes(x=TotalBsmtSF, y=lSalePrice))+geom_point()+geom_smooth(method="lm")
plot_grid(TotalBsmtSF,lTotalBsmtSF, labels = "AUTO")

#X1stFlrSF 
train_q2$lX1stFlrSF = log(train_q2$X1stFlrSF)
lX1stFlrSF <- train_q2%>%ggplot(aes(x=lX1stFlrSF, y=lSalePrice))+geom_point()+geom_smooth(method="lm")
X1stFlrSF <- train_q2%>%ggplot(aes(x=X1stFlrSF, y=lSalePrice))+geom_point()+geom_smooth(method="lm")
plot_grid(X1stFlrSF,lX1stFlrSF, labels = "AUTO")

#X2ndFlrSF
train_q2$lX2ndFlrSF = log(train_q2$X2ndFlrSF)
lX2ndFlrSF <- train_q2%>%ggplot(aes(x=lX2ndFlrSF, y=lSalePrice))+geom_point()+geom_smooth(method="lm")
X2ndFlrSF <- train_q2%>%ggplot(aes(x=X2ndFlrSF, y=lSalePrice))+geom_point()+geom_smooth(method="lm")
plot_grid(X2ndFlrSF,lX2ndFlrSF, labels = "AUTO")
```

#missing vars train_q2
```{r}
gg_miss_var(train_q2)
#lot frontage, garage year built, MasVnrArea
```

#remove lot frontage, garage year built, and masvnrarea
```{r}

train_q2 <- data.frame("Id"= train_q2$Id, "MSSubClass"= train_q2$MSSubClass, "LotArea"= train_q2$LotArea, "OverallQual"= train_q2$OverallQual, "OverallCond"= train_q2$OverallCond, "YearBuilt"= train_q2$YearBuilt, "YearRemodAdd"= train_q2$YearRemodAdd, "BsmtFinSF1"= train_q2$BsmtFinSF1, "BsmtFinSF2"= train_q2$BsmtFinSF2, "BsmtUnfSF"= train_q2$BsmtUnfSF, "TotalBsmtSF"= train_q2$TotalBsmtSF, "X1stFlrSF"= train_q2$X1stFlrSF, "X2ndFlrSF"= train_q2$X2ndFlrSF, "LowQualFinSF"= train_q2$LowQualFinSF, "BsmtFullBath"= train_q2$BsmtFullBath, "BsmtHalfBath"= train_q2$BsmtHalfBath, "FullBath"= train_q2$FullBath, "HalfBath"= train_q2$HalfBath, "BedroomAbvGr"= train_q2$BedroomAbvGr, "KitchenAbvGr"= train_q2$KitchenAbvGr, "TotRmsAbvGrd"= train_q2$TotRmsAbvGrd, "Fireplaces"= train_q2$Fireplaces, "GarageCars"= train_q2$GarageCars, "GarageArea"= train_q2$GarageArea, "WoodDeckSF"= train_q2$WoodDeckSF, "OpenPorchSF"= train_q2$OpenPorchSF, "EnclosedPorch"= train_q2$EnclosedPorch, "X3SsnPorch"= train_q2$X3SsnPorch, "ScreenPorch"= train_q2$ScreenPorch, "PoolArea"= train_q2$PoolArea, "MiscVal"= train_q2$MiscVal, "MoSold"= train_q2$MoSold, "YrSold"= train_q2$YrSold, "lGrLivArea"= train_q2$lGrLivArea, "lSalePrice"= train_q2$lSalePrice, "lX2ndFlrSF"= train_q2$lX2ndFlrSF, "lTotalBsmtSF"= train_q2$lTotalBsmtSF, "lX1stFlrSF"= train_q2$lX1stFlrSF)
```


#Error in lm.fit(x, y, offset = offset, singular.ok = singular.ok, ...) : NA/NaN/Inf in 'x'
```{r}
train_q2 <- NaRV.omit(train_q2)
```

#forward
```{r}
model_f <- lm(lSalePrice ~., data = train_q2)
f <-ols_step_forward_p(model_f, details=TRUE)
plot(f)
f
```

#backward
```{r}
model_b <- lm(lSalePrice ~ ., data = train_q2)
b <-ols_step_backward_p(model_b, details = TRUE)
plot(b)
b
```


#stepwwise
```{r}
model_s <- lm(lSalePrice ~ ., data = train_q2)
s <-ols_step_both_p(model_s, details = TRUE)
plot(s)
s
```


